# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r 'load'}
x <- read.csv("activity.csv")
x$date <- as.Date(x$date)
head(x)
```
## What is mean total number of steps taken per day?
```{r 'foo', echo=FALSE}

sumByDay <- function(x) {
  aggregate(steps ~ date, data=x, sum, na.action=na.omit)
}

head(sumByDay(x))

```
### histogram of the total number of steps taken each day
```{r 'hist unfilled'}

hist(sumByDay(x)$steps, breaks=24, xlab="Total Steps", main="Total Number of Steps Taken")
```
## What is the average daily activity pattern?
```{r 'daily stats'}
mean(sumByDay(x)$steps)
median(sumByDay(x)$steps)
```
### mean of values each day 
```{r 'daily mean'}

meanStepsByInterval <- function(x) {
  aggregate(steps ~ interval, data=x , mean, na.action=na.omit)
}

m <- meanStepsByInterval(x)

library(ggplot2)
qplot(m$interval, m$steps, geom="line", ylab="Ave. Steps", xlab="Interval")
```
# Interval with maximum number of steps:
```{r}
x[!is.na(x$steps) & x$steps == max(x$steps, na.rm=TRUE),]

```
# Imputing missing values
Missing values will be replaced by the average number of steps reported for that time interval.
```{r 'impute missing values'}
filledX <- x
filledX$isWeekday <- !weekdays(filledX$date, abbreviate=TRUE) %in% c("Sat", "Sun")

```
## Identify rows missing data
```{r 'id missing rows'}
missingRows <- which(is.na(filledX$steps))
```
## Define function to return the mean of a given interval across days
```{r 'interval mean function'}
intervalMean <- function(interval){
  sum(m[m$interval == interval,"steps"])
}
```
## Using sapply, replace missing values with the mean for that interval
```{r 'replace missing values'}
filledX[missingRows, "steps"] <- sapply(x[missingRows,]$interval, intervalMean)
hist(sumByDay(filledX)$steps, breaks=24,xlab="Total Steps", main="Total Number of Steps Taken (filled values)")
mean(sumByDay(filledX)$steps)
median(sumByDay(filledX)$steps)
```

# Differences in activity patterns between weekdays and weekends?
```{r 'weekend-weekday differences'}
filledX$isWeekday <- ifelse(weekdays(filledX$date, abbreviate=TRUE)
                            %in% c("Sat", "Sun"), "Weekend", "Weekday")

meanStepsByIntervalWkday <- function(x) {
  aggregate(steps ~ interval+isWeekday, data=x , mean, na.action=na.omit)
}

filledMeanX <- meanStepsByIntervalWkday(filledX)
qplot(interval, steps, data=filledMeanX, facets=isWeekday~., 
      geom="line", color=isWeekday) + stat_smooth(method="loess")
```